Emotion Recognition in Written Conversations Using CASTLE

### 2021-01-16 msccls/MScTemplateCS.tex
pierre comments taken care of

\documentclass[nofilelist]{cslthse-msc}
% to show a list of used packages at the end of the document, delete the nofilelist option
%\documentclass{cslthse-msc} 
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
%\usepackage{amsfonts}
%%\usepackage{amssymb}
\usepackage{amsthm}
%\usepackage{makeidx}
\usepackage{graphicx}
\usepackage[titletoc, header, page]{appendix}
\usepackage{transparent}
\usepackage{natbib}
\usepackage{wrapfig}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{subfig}
\usepackage{textcomp}
\usepackage{calligra}

@@

\acknowledgements{
If you want to thank people, do it here, on a separate right-hand page. Both the U.S. \textit{acknowledgments} and the British \textit{acknowledgements} spellings are acceptable.

We would like to thank Lennart Andersson for his feedback on this template.

We would also like thank Camilla Lekebjer for her contribution on this template, as well as Magnus Hultin for his popular science summary class and example document.

Thanks also go to the following (former) students for helping with feedback and suggestions on this template: Mikael Persson, Christoffer Lundgren, Mahmoud Nasser.
}

@@

%ML är automagiskt

\section{Background}
Natural language processing (NLP) is a field of study mainly in computer science that deals with analyzing human language in a way that computers can understand and process. According to \citet{ntlk2009}, natural language is used for everyday communication by humans, in contrast to artificial languages such as programming languages and mathematical notation. Since natural languages naturally evolve as time goes on and their adherence to structures diminish, it is difficult to define explicit rules to follow.

%prata om emotion recognition

%The LSTM architecture \citep{ntlk2009} is something....

%\begin{itemize}
%    \item Förklara vad systemet är
%    \item Berätta om företaget Telavox
%\end{itemize}

\section{Purpose}
Customer service is centered around the relation between company and customer with the goal of pleasing the customer. This goal can be very difficult to achieve and there is no straightforward way to do so. By analyzing prior conversations it is possible for companies to gain valuable insights in the most effective way to please the customer. Customer service is however a massive industry and it would take countless hours to analyze the conversations from even a single company manually. 
The company sponsoring this project approached us with an idea to try and solve this problem using machine learning. A usual technique for these types of problems is \textit{sentiment analysis} where utterances are labeled either as positive or negative. We instead decided to take another route, namely \textit{emotion recognition}. Emotion recognition focuses on labelling utterances from a non-binary set of emotions instead of sentiment analysis' binary method. This allows an analysis with greater specificity.   
In order to further the academic relevance of this project we decided that we would attempt to contribute to the field of emotion recognition by achieving state-of-the-art results.

\section{Telavox}
Telavox AB is a Swedish telecommunication company that specializes in business-to-business and business-to-customer unified communication as a service. As all of their customers deal in customer service they would all benefit from the solution previously discussed. Telavox therefore has a great interest in being able to offer this as a service. As Telavox has over 250 businesses as customers, the solution would facilitate the quality of customer service world wide.

\section{Research questions}
The primary aim of this thesis is to create a emotion recognition model that is able to produce state-of-the-art results. After this, a secondary aim is to investigate whether or not it is possible to create a dynamic model for instant response regarding customer service related conversations. The following questions are used as guidelines on how to realize these aims.

\begin{itemize}
    \item \textit{How does our model compare to other models in the same field evaluating on the same metrics?}
    \item \textit{Is it possible for an LSTM architecture to produce better results than a GRU one for emotion recognition?}
    \item \textit{After implementing a fully functioning model, how can we adapt it to predict utterances in real-time?}
\end{itemize}

@@

\section{Contributions}
%Potentialen av applications 
\section{Related work}

@@

The celebrity of computer science at this time is undoubtedly artificial intelligence (AI). A much more complex subject than most people know. Most of the complex problems in AI are solved using technologies within its sub-fields, which are illustrated in Figure \ref{fig:map}, rather than traditional artificial intelligence such as robotics \citep{bajwa_2020}. 
In order to solve most emotion recognition problems, there are two main concepts that are fundamental to build an accurate model. Firstly, the model should be created with deep learning algorithms. This enables it to find patterns that it otherwise could not be able to. Secondly, to effectively process the utterances it has been tasked to analyze, a word vector is often used to store the most commonly used words for easier analysis.

@@

\section{Machine Learning}
Machine learning \citep{franoischollet2017learning} is a field of study in artificial intelligence that aims to train a system to solve a problem, rather than to explicitly program it. By being presented with large amounts of data associated with the task at hand, the system algorithmically builds a statistical model that discovers structures in the data and allows the system to identify rules for the problem. 
In \textit{supervised learning} \citep{100pageBurkov}, which is one of the two main types of tasks in machine learning, the dataset is the collection of labeled examples $\{(\mathbf{x}_i, y_i)\}_{i=1}^N.$ Each element $\mathbf{x}_i$ is a feature vector that contains a value describing the example somehow and $y_i$ is a label, which is either a finite set of classes or a real number. The purpose of supervised learning is to use the dataset to produce a model that takes a feature vector x as input and outputs information that allows deducing the label y for this feature vector.) 
The other main type of task in machine learning is \textit{unsupervised learning}. The dataset for unsupervised learning is $\{ (\mathbf{x}_i)\}_{i=1}^N$ where $\mathbf{x}_i$ is a feature vector. The goal of unsupervised learning is to use the dataset to produce a model that takes a feature vector x as input and outputs either another vector or a value.

@@

Neural networks are machine learning systems loosely inspired by biological neural networks. A neural network is a collection of connected nodes or neurons that was first mentioned in \citet{mcculloch1943}. A neuron \citep{dawson1998ann} is an information-processing unit that receives inputs and produces a single output that can be sent to multiple other neurons. To find the output of a neuron the weighted sum of all inputs is weighted by the weights all connecting to the neuron. This weighted sum is then passed through an activation function to provide the output.

@@

\begin{figure}[htp]
    \centering
    \includegraphics[width=12cm]{msccls/explanatory_images/single_neuron.png}
    \caption{An example of a single neuron.}
    \label{fig:neuron}
\end{figure}

@@

Neural networks can in turn be structured together and form multiple layers that progressively learn. These networks fall under the deep learning umbrella. Deep learning has several closely relevant definitions, where \citet{deng2014deep} mentions the following:

\begin{quote}{A class of machine learning techniques that exploit many layers of non-linear information processing for supervised or unsupervised feature extraction and transformation, and for pattern analysis and classification}
\end{quote}

Each of the layers in a deep learning network manipulates their input data in some way, and the result is stored in the layer's weights. The results are monitored with the help of the \textit{loss function}. The loss function returns a distance score that indicates how far off the network's previous prediction was from the true values \citep{franoischollet2017learning}. The process of reducing the distance score that the loss function produces is called training. 
Using the score produced by the loss function as feedback, the \textit{backpropagation} algorithm computes the \textit{gradient}, a multi-variable generalization of the derivative, which is then used to adjust the weights in such a way that the distance score produced by the loss function is minimized.

@@

One of the simpler neural network classes used in deep learning is recurrent neural networks (RNN) and they process either sequential data or time-series data \citep{lipton2015critical}. Since the inputs are sequential, all the inputs are dependant on each other in one way or another. RNNs are therefore suitable for processing sequences of text, audio, and video where every computation will be performed on each element in the sequence.

One of RNN's weaknesses is that they suffer the \textit{vanishing gradient problem} \citep{hochreiter1998}. Theoretically, RNNs can process arbitrarily long sequences, in practice, they are limited to the length of the sequence. If the sequence is too long, the network will struggle to try to look back in time longer than a few steps away. In a lot of cases, this will result in leaving out important information from the prediction. There are instead other alternatives that consider this weakness.

@@

 \begin{figure}[!ht]
    \centering
    \includegraphics[scale=0.3]{msccls/explanatory_images/rnn.png}
    \caption{A simple RNN node}
    \label{fig:rnn_node}
\end{figure}

\subsection{Gated Recurrent Unit -- GRU}
%Byta alla "node" till "cell" för att lättare kunna referera till LSTM-celler i CASTLE
A gated recurrent unit (GRU) is recurrent neural network architecture introduced by \citet{cho2014learning} and is considered an improvement to the traditional recurrent neural networks. Basic GRU nodes  are primarily constructed of an \textit{update gate} and a \textit{reset gate}. The update gate regulates the amount of information that is passed along to the next state, while the reset gate on the other hand decides the amount of past information that is going to be disregarded henceforth. 
Although fairly new it is already a well-established variant of recurrent neural networks. In comparison to LSTM layers, GRU layers are cheaper to run since they use fewer parameters, but have less representational power. It has been shown that GRU layers have had better performance on smaller and less frequent datasets than LSTM layers \citep{Gruber2020AreGC}.

@@

\subsection{Long Short-Term Memory -- LSTM}
Another widely used recurrent neural network architecture is Long Short-Term Memory (LSTM). It was proposed by \citet{hochreiter1997} as a solution to the vanishing gradient problem that might occur in recurrent neural networks and has barely been changed since. The vanishing gradient problem occurs in machine learning when the gradient progressively decreases, which prevents the adjustments of the weight's values.

As illustrated in \ref{fig:lstm_node} a LSTM node consists of an \textit{input gate}, an \textit{output gate}, a \textit{forget gate} and a \textit{cell state}. 
\begin{itemize}
    \item The input gate controls the extent of data to write onto the internal cell-state.
    \item The output gate controls the extent of data to pass as output. 
    \item The forget gate, similarly to the GRUs reset gate, decides how much of the previous data is irrelevant and will be ignored. 
    \item The cell state stores the previous state of a node and passes it on to its following node. 
\end{itemize}

\begin{figure}[!ht]
    \centering
    \includegraphics[scale=0.45]{msccls/explanatory_images/lstm.png}
    \caption{A LSTM node}
    \label{fig:lstm_node}
\end{figure}

\section{Training Process}
In order to train a model there must exist some data set, either labeled or unlabeled depending on whether supervised or unsupervised learning is the targeted method, that feeds the model with input. When using a labeled data set the labels are regarded as the solution to the problem. When training a model the goal is to find values for its weights such that the model correctly labels the given input. The training of a model is done step by step in the shape of \textit{epochs}. Epochs represent the number of iterations that the model has passed through a data set intended for training the model. Following each epoch, a \textit{loss function} measures how bad the model's previous prediction was and produces a value for the \textit{loss}. The model will have a perfect prediction when the loss reaches zero, therefore minimizing the loss is desirable.

In order to minimize the loss you need to know what the best values for the weights to do so are. To do this, while training the model iteratively calculates the derivatives of the

We want to travel through the function and find the place where the derivative is zero. This is done for each weight in the network. It is accomplished by iteratively using backpropagation algorithm that finds a value closer the the derivative being zero. Gradient descent is then used to update the weights to match the result of the backpropagation algorithm.

direction the weights are taking the model. This is done by evaluating the derivatives of the ...in the network and that is exactly what the \textit{gradient descent} does.

@@

@@

\section{Evaluation}
Many different metrics can be used to evaluate a machine learning model, based on what kind of model and what its purpose is. For classification models, many of the metrics are based on a confusion matrix \citep{FAWCETT2006861}. In a confusion matrix, one of the axes represents the actual class of the instance the model is trying to predict and the other represents what instance the model predicted. The most basic and common confusion matrix is a binary $2 \cdot 2$ matrix, as seen in Figure \ref{fig:confusion}. In binary cases, the instances usually can be represented as a yes or no question, such as \textit{is there a bird in this picture?}

@@

For cases where we have multi-class classification instead of binary, such as when determining emotions, the same logic and matrix is used but on a bigger scale and with the specific classes instead of positive and negative. The equations provided in subsequent sections to display how to calculate metrics are identical for multi-class context, but the classification instead represents the sum of the classification for each class. For instance, when determining if an emotion is \textit{Neutral}, \textit{Happy}, or \textit{Angry}, $TP_{neutral}$ would represent when it got neutral correct, and TP would be the sum of all emotion predictions it got correct.

@@

\subsection{Accuracy}
One of the simplest metrics used, accuracy quantifies how many percent of predictions that the model got right and is calculated by taking all instances where our model predicted positive and dividing it with all instances.

@@

If we have a balanced data set with a similar amount of positive and negative instances, this can give valuable information on the performance of the model since getting predictions right is the goal of any model. If, however, our set is imbalanced and contains e.g. 90 percent negative instances and only 10 percent positive, the model can achieve a 90 percent accuracy by simply predicting negative on every instance so the accuracy suddenly becomes useless \citep{imbalancedlearning}.

@@

$$ Precision = \frac{TP}{TP + FP}$$

Precision gives us information on how accurate the model is when predicting an instance to be true. This is useful when dealing with instances where it's important to not give a false positive. If we have a model that is trying to predict whether or not a suspect is guilty of a certain crime, a high precision would make it unlikely for an innocent person to be determined as a criminal. 
However, precision does not give any information on the number of false negatives present, meaning that a high precision could mean that the model seldom identifies a person, but when it does identify someone, that said person is likely guilty.

@@

$$ Recall =\frac{TP}{TP + FN}$$

As a sort of counterpoint to precision which tells us how good the model is when guessing true, recall tells us how many positive instances the model missed. A high recall can be extremely important when dealing with subjects such as contagious diseases where a result of false-negative can be devastating. Recall doesn't take into account how many positive predictions turned out to be false so in cases where that doesn't matter and you want to predict positive at the slightest probability of it being true, optimizing for a high recall is essential.

\subsection{F1-score}
A combination of the two previous metrics, F1-score is calculated by creating the harmonic mean of \textit{precision} and \textit{recall}.

$$ F1score = 2 \cdot \frac{Precision \cdot Recall}{Precision + Recall} $$

In cases where there is no extra importance in finding either false negatives or false positives, you just want to minimize the risk of errors whatsoever. The harmonic mean is a kind of average that is appropriate when working with ratios, which is what both precision and recall are. This lets us improve the model concerning either of the variables and not have to think about if, for example, the reduction in Recall was worth the gain in Precision. In multi-class scenarios with imbalanced data sets there are multiple ways to calculate the F1-score, namely \textit{macro}, \textit{weighted} and \textit{micro}. Which one to use depends on your data set and what you prioritize in your model.

The macro F1-score is derived by calculating the F1-score for each class separately and then just summing them up, leading to disproportionate weight on the classes with fewer instances. If we assume that we are using a data set where the \textit{neutral} class is represented in 10 000 instances and a less common emotion \textit{fear} in just 50. The model then managed to predict 9 995 of the neutral instances, but only 25 of the fear ones. When calculating the mean F1-score, the low fear score would muddy the entire score and make it seem bad, when in fact the score for the neutral case is excellent. If we want to optimize for the minority classes, however, this lets us do just that.

@@

When calculating micro F1-score, instead of calculating the f1-scores separately, the f1-score is calculated directly by using the total values for TP, FN, and FP in the Precision and Recall formulas which places no bias on either class. If we have an example data set with two minority classes \textit{anger} and \textit{joy}, we can see that for each instance where anger is true, classifying it as joy would be an FN from the perspective of the anger class. From the perspective of the joy class, however, it would instead be an FP. Since micro calculations take into account the total values for our identification labels and each FN correspond to the FP of another class, $FP = FN$, which in turn means that $Precision = Recall$ in these cases. Precision and Recall being equal means that their mean also is equal and thus \textit{F1-}$score_{micro} = Precision_{Micro} = Recall_{Micro}$ which incidentally also equals \textit{Accuracy} as this is the proportion of instances we got correct. This means that just like Accuracy this is a good metric on balanced data sets but can be deceiving in an imbalanced one.

\section{Loss}
%might be removed

\section{Overfitting}
Overfitting is a concept in statistical analysis in which the result corresponds too closely to a certain data set. Instead of the model learning how to notice and predict features, it simply remembers a massive amount of examples from the training set. This causes the model to perform very well on that specific data but would likely have a very high error rate on any new data.

\section{Natural Language Processing -- NLP}
Natural languages are the set of languages that humans have evolved naturally without any predetermined strategy like e.g English or Mandarin. Making computers understand and analyze natural languages has long been a challenge and one of the first practical attempts was Weaver and Booth's translation project in 1946 trying to crack enemy codes during World War II \citep{Liddy2001NLP}. Shortly thereafter \citet{Turing1950} created his famous \textit{Imitation game} or more commonly referred to as the Turing test, which has developed into a test to evaluate whether a computer's natural language response in a real-time written conversation can be indistinguishable from that of a human. One of the reasons that make natural language processing a challenge is the fact that natural languages are minted with unstructured types of data such as inflection or dialects, or using phrases to convey the opposite than the common meaning namely sarcasm. Natural languages are also oftentimes ambiguous and require a context-based interpretation that computers have a hard time understanding.

@@

\section{CASTLE}

>>>>

\documentclass[nofilelist]{cslthse-msc}
% to show a list of used packages at the end of the document, delete the nofilelist option
%\documentclass{cslthse-msc} 
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
%\usepackage{amsfonts}
%%\usepackage{amssymb}
\usepackage{amsthm}
%\usepackage{makeidx}
\usepackage{graphicx}
\usepackage[titletoc, header, page]{appendix}
\usepackage{transparent}
\usepackage{natbib}
\usepackage{wrapfig}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{subfig}
\usepackage{textcomp}
\usepackage{calligra}
\usepackage{nameref}

@@

\acknowledgements{

We would like express our gratitude, first and foremost, to our supervisor Pierre. There is no way that this project would have been completed without your guidance, at least not in time.

We would also like to thank \textit{Telavox} and our company supervisor Sara for our weekly meetings, where much of our planning and reflection were made.

As the COVID-19 epidemic created some unusual circumstances, we would like to give a special thanks to \textit{Åke P Fågel \& Vilt} for providing us with a stable office to work out of. 
}

@@

%ML är automagiskt - PopSci

\section{Background}
Natural language processing (NLP) is a field of study mainly in computer science that deals with analyzing human language in a way that computers can understand and process. According to \citet{ntlk2009}, natural language is used for everyday communication by humans, in contrast to artificial languages such as programming languages and mathematical notation. Since natural languages naturally evolve as time goes on and their adherence to structure diminish, it is difficult to define explicit rules to follow. For this reason, there is a need for machine learning models to aid the computer in identifying these diffuse rules.

In order to feed the model with effective information, the input is structured and inserted into a data set. This way, the computer has an easier way of processing and analyzing the data. When handling data sets constructed using exclusively textual input, the data sets are instead referred to as \textit{corpora}.

%?? and hence the need of machine learning to replace rules?? ?? You need also to introduce more specifically your task ??

\section{Purpose}
Customer service is centered around the relation between company and customer with the goal of pleasing the customer. This goal can be very difficult to achieve and there is no straightforward way to do so. By analyzing prior conversations, it is possible for companies to gain valuable insights in the most effective way to please the customer. Customer service is however a massive industry and it would take countless hours to analyze the conversations from even a single company manually.

The company sponsoring this project approached us with an idea to try and solve this problem using machine learning. A usual technique for these types of problems is \textit{sentiment analysis}, where utterances are labeled either as positive or negative. We instead decided to take another route, namely \textit{emotion recognition}. Emotion recognition focuses on labelling utterances from a non-binary set of emotions instead of sentiment analysis' binary method. This allows an analysis with greater specificity.

In order to further the academic relevance of this project we decided that we would attempt to contribute to the field of emotion recognition by achieving state-of-the-art results.

\section{State-of-the-art}
In this paper, the term state-of-the-art will be recurringly mentioned as it is a fundamental part of our aim and result. The term refers to the highest level of achievement accomplished. A common goal in machine learning projects is to attempt to raise the bar for the current state-of-the-art results across various metrics for the task that the project is trying to solve. Presenting state-of-the-art results is an excellent approach in proving that an implementation or algorithm is valid and valuable.

%Description of what it is and why it plays a big part in advancing algorithms and models. Deepminds Alphafold has led to groundbreaking results in a competitive environment.

\section{Telavox}
We carried out this project at Telavox AB. Telavox is a Swedish telecommunication company that specializes in business-to-business and business-to-customer unified communication as a service. As all of their customers deal in customer service, they would all benefit from the solution previously discussed. Telavox therefore has a great interest in being able to offer this as a service. As Telavox has over 250 businesses as customers, the solution would facilitate the quality of customer service world wide.

\section{Research questions}
\label{research_questions}
The primary aim of this thesis is to create a emotion recognition model that is able to produce state-of-the-art results. After this, a secondary aim is to investigate whether or not it is possible to create a dynamic model for instant response regarding customer service related conversations. The following questions are used as guidelines on how to realize these aims.

\begin{itemize}
    \item \textit{Is it possible to create a model that fulfills our aims?}
%?? THe first question is can we create such a model??
    \item \textit{Could a different architecture than the current state-of-the-art produce better results?}
%?? here your statement is too specific as you did not introduce the terms. Use vague words such as What are the different option to design an architecture and train a model?? 
%?? I have swapped the items ??
    \item \textit{How does our model compare to other models in the same field evaluating on the same metrics?}
    \item \textit{After implementing a fully functioning model, how can we adapt it to predict utterances in real-time?}
    %Snacka om detta i limitations
\end{itemize}

@@

@@

The celebrity of computer science at this time is undoubtedly \emph{artificial intelligence} (AI). A much more complex subject than most people know. Most of the complex problems in AI are solved using technologies within its sub-fields, which are illustrated in Figure \ref{fig:map}, rather than traditional artificial intelligence such as robotics \citep{bajwa_2020}.

%In order to solve most emotion recognition problems, there are two main concepts that are fundamental to build an accurate model. Firstly, the model should be created with deep learning algorithms. This enables it to find patterns that it otherwise could not be able to. Secondly, to effectively process the utterances it has been tasked to analyze, a word vector is often used to store the most commonly used words for easier analysis. ?? I propose this instead??

In order to solve most emotion recognition problems, there are two main concepts that are fundamental to build an accurate model:
\begin{enumerate}
\item As shown by the recent progress in the field \citep{deep_survey}, the model should be created with deep learning algorithms. This should enable it to find patterns that it otherwise could not be able to. 
\item To effectively process the utterances, deep learning requires to collect an annotated corpus and design an appropriate representation for the words.
\end{enumerate}

@@

\section{Machine Learning}
Machine learning \citep{franoischollet2017learning} is a field of study in artificial intelligence that aims to train a system to solve a problem, rather than to explicitly program the system. By being presented with large amounts of data associated with the task at hand, the system algorithmically builds a statistical model that discovers structures in the data and allows the system to identify rules for the problem. 
In \textit{supervised learning} \citep{100pageBurkov}, which is one of the two main types of tasks in machine learning, the dataset is the collection of labeled examples $\{(\mathbf{x}_i, y_i)\}_{i=1}^N.$ Each element $\mathbf{x}_i$ is a feature vector that contains a value describing the example somehow and $y_i$ is a label, which is either a finite set of classes or a real number. The purpose of supervised learning is to use the dataset to produce a model that takes a feature vector $\mathbf{x}$ as input and outputs information that allows deducing the label $y$ for this feature vector.)

The other main type of task in machine learning is \textit{unsupervised learning}. The dataset for unsupervised learning is $\{ (\mathbf{x}_i)\}_{i=1}^N$ where $\mathbf{x}_i$ is a feature vector. The goal of unsupervised learning is to use the dataset to produce a model that takes a feature vector $\mathbf{x}$ as input and outputs either another vector or a value.

@@

Neural networks are machine learning systems loosely inspired by biological neural networks. A neural network is a collection of connected nodes or neurons that was first mentioned in \citet{mcculloch1943}. A neuron \citep{dawson1998ann} is an information-processing unit that receives multiple inputs and produces a single output that can be sent to multiple other neurons. To find the output of a neuron the weighted sum of all inputs is weighted by the weights all connecting to the neuron. This weighted sum is then passed through an activation function to provide the output.

@@

\begin{figure}[htp]
    \centering
    \includegraphics[width=12cm]{msccls/explanatory_images/neuron_new.png}
    \caption{An example of a single neuron. From \citet{wiki:neuron_image}.}
    \label{fig:neuron}
\end{figure}

@@

Neural networks can in turn be structured together and form multiple layers that progressively learn. These networks fall under the deep learning umbrella. Deep learning has several closely related definitions, where \citet{deng2014deep} mentions the following:

\begin{quote}{A class of machine learning techniques that exploit many layers of non-linear information processing for supervised or unsupervised feature extraction and transformation, and for pattern analysis and classification.}
\end{quote}

Each of the layers in a deep learning network manipulates its input data in some way, and the result is stored in the layer's weights. The results are monitored with the help of the \textit{loss function}. The loss function returns a distance score that indicates how far off the network's previous prediction was from the true values \citep{franoischollet2017learning}. The concept of reducing the distance score that the loss function produces is called training.

Using the score produced by the loss function as feedback, the \textit{backpropagation} algorithm computes the \textit{gradient}, a multi-variable generalization of the derivative, which is then used to adjust the weights in such a way that the distance score produced by the loss function is minimized.

@@

One of the simplest neural network classes used in deep learning is \emph{recurrent neural networks} (RNN). RNN process either sequential data or time-series data \citep{lipton2015critical}. Since the inputs are sequential, all the inputs are dependent on each other in one way or another. RNNs are therefore suitable for processing sequences of text, audio, and video where every computation will be performed on each element in the sequence.

One of RNN's weaknesses is that they suffer from the \textit{vanishing gradient problem} \citep{hochreiter1998}. While RNN's theoretically can process arbitrarily long sequences, in practice, they are limited to the length of the sequence. If the sequence is too long, the network will struggle to try to look back in time longer than a few steps away. In a lot of cases, this will result in leaving out important information from the prediction. There are instead other alternatives that consider this weakness.

@@

 \begin{figure}[!ht]
    \centering
    \includegraphics[scale=0.5]{msccls/explanatory_images/rnn.png}
    \caption{A simple RNN node}
    \label{fig:rnn_node}
\end{figure}

\subsection{Long Short-Term Memory -- LSTM}
%?? I would have this section before GRU ??

Another widely used recurrent neural network architecture is Long Short-Term Memory (LSTM). It was proposed by \citet{hochreiter1997} as a solution to the vanishing gradient problem that might occur in recurrent neural networks and has barely been changed since. The vanishing gradient problem occurs in machine learning when the gradient progressively decreases, which prevents the adjustments of the weight's values.

As illustrated in Figure \ref{fig:lstm_node}, a LSTM node consists of an \textit{input gate}, an \textit{output gate}, a \textit{forget gate} and a \textit{cell state}. 
\begin{itemize}
    \item The input gate controls the extent of data to write onto the internal cell-state.
    \item The output gate controls the extent of data to pass as output. 
    \item The forget gate that decides how much of the previous data is irrelevant and will be ignored. 
    \item The cell state stores the previous state of a node and passes it on to its following node. 
\end{itemize}

\begin{figure}[!ht]
    \centering
    \includegraphics[scale=0.5]{msccls/explanatory_images/lstm.png}
    \caption{A LSTM node} 
    %?? You need to mention the source of your figure if it is not yours??
    \label{fig:lstm_node}
\end{figure}

\subsection{Gated Recurrent Unit -- GRU}
%Byta alla "node" till "cell" för att lättare kunna referera till LSTM-celler i CASTLE
A \emph{gated recurrent unit} (GRU) is recurrent neural network architecture introduced by \citet{cho2014learning} and is considered an improvement to the traditional recurrent neural networks.

Basic GRU nodes  are primarily constructed of an \textit{update gate} and a \textit{reset gate}. The update gate regulates the amount of information that is passed along to the next state, while the reset gate, much like the forget gate of LSTM, decides the amount of past information that is going to be disregarded henceforth.

Although fairly new, it is already a well-established variant of recurrent neural networks. In comparison to \textit{LSTM} layers, GRU layers are cheaper to run since they use fewer parameters, but have less representational power. It has been shown by \citet{Gruber2020AreGC} that GRU layers have had better performance on smaller and less frequent datasets than LSTM layers.

%?? You cannot cite LSTM without describing them first. In addition GRU are newer than LSTMs??

@@

\section{Training Process}
In order to train a model, there must exist some data set, either labeled or unlabeled depending on whether supervised or unsupervised learning is the targeted method, that feeds the model with input. When using a labeled data set, the labels are regarded as the solution to the problem. When training a model, the goal is to find values for its weights such that the model correctly labels the given input.

The training of a model is done step by step in the form of \textit{epochs}. Epochs represent the number of times that the model iterates through a data set that is intended to train the model. Following each epoch, a \textit{loss function} measures how incorrect the model's previous prediction was and produces a value for the \textit{loss}. The model will have a perfect prediction when the loss reaches zero, therefore minimizing the loss is desirable.

The \textit{backpropagation} algorithm is a fundamental tool for minimizing the loss. The algorithm starts with the loss and works its way \textit{backwards} through the layers of the network and applies the chain rule to determine how much impact each of the weights has had on the loss. This impact is then used for the \textit{gradient descent} algorithm to alter the weights in the direction of flattening out the derivative. While travelling through the function, the ultimate goal is to find where the derivative is zero and this is done for each weight in the network. Gradient descent updates the weights to match the results from the backpropagation algorithm.

%bild på training process

%Man får inte minimera lossen för aggressivt, det finns en balans, annars sker overfitting vilket är......

%Overfitting is a concept in statistical analysis in which the result corresponds too closely to a certain data set. ?? This is not the definition of overfitting. Overfitting is when the model fits well the training but fails outside it?? Instead of the model learning how to notice and predict features, it simply remembers a massive amount of examples from the training set. This causes the model to perform very well on that specific data but would likely have a very high error rate on any new data.

%bild på overfitting

%We want to travel through the function and find the place where the derivative is zero. This is done for each weight in the network. It is accomplished by iteratively using backpropagation algorithm that finds a value closer the the derivative being zero. Gradient descent is then used to update the weights to match the result of the backpropagation algorithm.

%In order to minimize the loss you need to know what the best values for the weights to do so are. To do this, while training the model iteratively calculates the derivatives of the

%We want to travel through the function and find the place where the derivative is zero. This is done for each weight in the network. It is accomplished by iteratively using backpropagation algorithm that finds a value closer the the derivative being zero. Gradient descent is then used to update the weights to match the result of the backpropagation algorithm.

%Backpropagation, givet en loss och weights, så tar reda på hur mcyekt varje weight har påverkat lossen. Gradient decent använder denna infon för att uppdatera sina vikter och gör om

%direction the weights are taking the model. This is done by evaluating the derivatives of the ...in the network and that is exactly what the \textit{gradient descent} does.

%so all you have to do is find all the points where the derivative goes to 0 and check for which of these points the function has the lowest value. The engine of neural networks: gradient-based optimization. Applied to a neural network, that means finding analytically the combination of weight values that yields the smallest possible loss function.

@@

%Gradient Descent is an optimization technique that is used to find the parameters that will give the least loss.

@@

\section{Evaluation}
Many different metrics can be used to evaluate a machine learning model, based on what kind of model and what its purpose is. For classification models, many of the metrics are based on a confusion matrix \citep{FAWCETT2006861}.

In a confusion matrix, one of the axes represents the actual class of the instance the model is trying to predict and the other represents what instance the model predicted. The most basic and common confusion matrix is a binary $2 \cdot 2$ matrix, as seen in Figure \ref{fig:confusion}. In binary cases, the instances usually can be represented as a yes or no question, such as \textit{is there a bird in this picture?}

@@

For cases where we have multi-class classification instead of binary, such as when determining emotions, the same logic and matrix is used but on a bigger scale and with the specific classes instead of positive and negative.

The equations provided in subsequent sections to display how to calculate metrics are identical for multi-class context, but the classification instead represents the sum of the classification for each class. For instance, when determining if an emotion is \textit{Neutral}, \textit{Happy}, or \textit{Angry}, $TP_{neutral}$ would represent when it got neutral correct, and TP would be the sum of all emotion predictions it got correct.

@@

\subsection{Accuracy}
Accuracy quantifies how many percent of predictions the model got right and is calculated by taking all instances, where our model predicted positive and dividing it with all instances:

@@

If we have a balanced dataset with a similar amount of positive and negative instances, then this can give valuable information on the performance of the model since getting predictions right is the goal of any model. If, however, our set is imbalanced and contains e.g. 90 percent negative instances and only 10 percent positive, the model can achieve a 90 percent accuracy by simply predicting negative on every instance so the accuracy suddenly becomes useless \citep{imbalancedlearning}.

@@

$$ Precision = \frac{TP}{TP + FP}.$$

Precision gives us information on how correct the model is when predicting an instance to be true. This is useful when dealing with instances, where it's important to not give a false positive. If we have a model that is trying to predict whether or not a suspect is guilty of a certain crime, a high precision would make it unlikely for an innocent person to be determined as a criminal.

However, precision does not give any information on the number of false negatives present, meaning that a high precision could mean that the model seldom identifies a person, but when it does identify someone, that said person is likely guilty.

@@

$$ Recall =\frac{TP}{TP + FN}.$$

As a sort of counterpoint to precision which tells us how good the model is when guessing true, recall tells us how many positive instances the model missed. A high recall can be extremely important when dealing with subjects such as contagious diseases where a result of false-negative can be devastating.

Recall does not take into account how many positive predictions turned out to be false so in cases, where that does not matter and you want to predict positive at the slightest probability of it being true, optimizing for a high recall is essential.

\subsection{F1 score}
A combination of the two previous metrics, the F1 score, is calculated by creating the harmonic mean of precision and recall:

$$ F1score = 2 \cdot \frac{Precision \cdot Recall}{Precision + Recall}.$$

In cases where there is no extra cost associated with either false negatives or false positives, the goal is usually to minimize the total amount of errors. The harmonic mean is a type of average that is appropriate when working with ratios, which is what both precision and recall are.

This lets us improve the model concerning either of the variables and not have to think about if, for example, the reduction in Recall was worth the gain in Precision. In multi-class scenarios with imbalanced data sets there are multiple ways to calculate the F1-score, namely \textit{macro}, \textit{weighted} and \textit{micro}. Which one to use depends on your data set and what you prioritize in your model.

The macro F1-score is derived by calculating the F1-score for each class separately and then just summing them up, leading to an excessive focus on the classes with fewer instances. If we assume that we are using a data set where the \textit{neutral} class is represented in 10 000 instances and a less common emotion \textit{fear} in just 50. The model then managed to predict 9 995 of the neutral instances, but only 25 of the fear ones. When calculating the mean F1-score, the low fear score would muddy the entire score and make it seem bad, when in fact the score for the neutral case is excellent. If we want to optimize for the minority classes, however, this lets us do just that.

@@

When calculating micro F1 score, instead of calculating the F1 scores separately, the F1 score is calculated directly by using the total values for TP, FN, and FP in the Precision and Recall formulas which places no bias on either class. If we have an example data set with two minority classes \textit{anger} and \textit{joy}, we can see that for each instance where anger is true, classifying it as joy would be an FN from the perspective of the anger class. From the perspective of the joy class, however, it would instead be an FP. Since micro calculations take into account the total values for our identification labels and each FN correspond to the FP of another class, $FP = FN$, which in turn means that $Precision = Recall$ in these cases.

Precision and Recall being equal means that their mean also is equal and thus \textit{F1-}$score_{micro} = Precision_{Micro} = Recall_{Micro}$ which incidentally also equals \textit{Accuracy} as this is the proportion of instances we got correct. This means that just like Accuracy, this is a good metric on balanced data sets but can be deceiving in an imbalanced one.

\section{Natural Language Processing -- NLP}
Natural languages are the set of languages that humans have evolved naturally without any predetermined strategy such as English or Mandarin. Making computers understand and analyze natural languages has long been a challenge and one of the first practical attempts was Weaver and Booth's translation project in 1946 trying to crack enemy codes during World War II \citep{Liddy2001NLP}. Shortly thereafter \citet{Turing1950} created his famous \textit{Imitation game}, or more commonly referred to as the Turing test, which has developed into a test to evaluate whether a computer's natural language response in a real-time written conversation can be indistinguishable from that of a human.

One of the reasons that make natural language processing a challenge is the fact that natural languages are minted with unstructured types of data such as inflection or dialects, or using phrases to convey the opposite than the common meaning namely sarcasm. Natural languages are also oftentimes ambiguous and require a context-based interpretation that computers have a hard time understanding.

@@

%?? You could write a bit more on COSMIC ??
